\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-02-17 forwards-and-backwards.tex
alison edits 2/15/23

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR^d})} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on space x level

@@

It is of interest to understand under what conditions we can replace the two-step limiting
process described above by one in which we simultaneously scale the 
kernels
and the other parameters in 
our population model to arrive at the PDE limit.
This is mathematically much more challenging, but we establish such 
one-step convergence in cases for which the limit is a classical 
reaction diffusion equation of the form~(\ref{semilinear PDE}) with $\DG=\Delta$
and $\rho_F$ is a Gaussian density. We allow a wide class of reaction terms, $F$, so
that the Fisher-KPP equation emerges as a special case.

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity that have been laid down; from 
genetic data, one can try to infer the genealogical trees that relate 
individuals in a sample from the population, and these have been shaped 
by its history. It is therefore of interest
to establish information about the distribution of genealogical trees 
under our population model.
To achieve this we exploit a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a huge range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2018}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

We aim to capture universal behaviour by passing to a scaling limit. Specifically, 
we shall take the ``density'', $N$, to infinity,
and also scale time by a factor of $\theta$,
in such a way that defining $\eta^N(t) = X(\theta t) / N$,
the process $\{\eta^N(t)\}_{t \ge 0}$
will converge to a suitable measure-valued process
as $N$ and $\theta$ tend to infinity,
with the nature of the limit depending on how they tend to infinity together.
Evidently, we shall also need to scale the dispersal kernel if we are to obtain a
nontrivial limit. We write $q_\theta(x,\cdot)$ for the density of the 
multivariate Gaussian obtained by
multiplying the mean and variance components corresponding to $q(x,\cdot)$ by $1/\theta$.

@@

\begin{remark}
Although this model allows fairly general birth and death mechanisms,
there are a number of limitations.
Perhaps most obviously, to simplify the notation
individuals give birth to only one offspring at a time,
although this restriction could be easily lifted
\citep[as in Section 3.4 of][]{etheridge/kurtz:2018}.
Furthermore, individuals do not move during their lifetime,
and the age of an individual does not affect its fecundity or death rate.
Finally, there is no notion of mating
(although limitations on reproduction due to availability of mates can be incorporated into the birth rate, $\gamma$),
so the lineages we follow will be uniparental.
For these reasons, the model is most obviously applicable to bacterial populations or selfing plants,
although we do not anticipate that incorporation of these complications
will change the general picture.
\end{remark}

@@

\begin{notation}
Expressions like $\gamma(x, \smooth{\gamma} \eta(x))$ will appear repeatedly in what follows.
To make formulae more readable, we overload notation to define
$$
    \gamma(x, \eta) := \gamma(x, \smooth{\gamma} \eta(x)) ,
$$
and similarly write $r(x, \eta)$ for $r(x, \smooth{r} \eta(x))$
and $\mu_\theta(x, \eta)$ for the expression of equation \eqref{eqn:mu_defn}.
When convenient, we may also suppress the arguments completely,
writing simply $\gamma$, $r$, and $\mu_\theta$ for these quantities.
\end{notation}

@@

In summary, at each time $t$, $\eta^N_t$ is purely atomic, consisting of atoms of mass $1/N$.
At instantaneous rate $\theta \gamma(x, \eta^N_t) N \eta^N_t(dx)$ an offspring of mass $1/N$ is 
produced at location $x$; it disperses to 
a location $y$ offset from $x$
by an independent Gaussian with mean $\meanq(x) / \theta$
and covariance matrix $\covq(x) / \theta$, and once there establishes instantaneously
with probability $r(y, \eta^N_t)$, or else dies. 
At instantaneous rate $\theta \mu_\theta(x, \eta^N_t) N \eta^N_t(dx)$ an individual at location 
$x$ dies.
Note that the process 
$\left(\eta^N_t\right)_{t\ge 0}$, which
records numbers and locations of adult individuals, is just a scaled spatial birth and 
death process. If, for example, we insist that $\gamma(x,m)$ is bounded, then 
existence (and in particular non-explosion) is guaranteed by comparison with
a pure birth process. We do not dwell on this, as we shall require more stringent conditions
if we are to pass to the limit as $\theta$ and $N$ tend to infinity.

It is convenient to characterise the process as a solution to a martingale problem.
We write $C_b^\infty(\IR^d)$ for the space of bounded smooth functions on $\IR^d$, and, where
convenient, we
write $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$.
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction}
	For each value of $N$ and $\theta$, and each purely atomic 
$\eta_0^N\in\measures$ with atoms of mass $1/N$,
$(\eta^N_t)_{t \geq 0}$ is the (scaled) empirical measure of a birth-death 
process with c\`adl\`ag paths in $\measures$ for which,
for all $f \in C^{\infty}_{b}(\IR^d)$,
writing $q_\theta(x, dy)$ for the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
\\ & \qquad \qquad \qquad \qquad {}
        \times \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
%\\ & \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale (with respect to the natural filtration), 
with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N(f)]_t =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \Big\langle \gamma(x, \eta^N_{s})
        \int f^2(z)r(z, \eta^N_{s}) q_\theta(x,dz) 
    , \eta^N_{s}(dx) \Big\rangle \\
& \qquad \qquad {}
    + \Big\langle \mu_\theta(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\Big\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}
The form of~(\ref{eqn:eta_martingale}) and~(\ref{eqn:prelimit_martingale_variation})
is explained in Section~\ref{sec:heuristics}.
Note that since individuals are produced at rate $N \gamma \eta$, but each
has mass $1/N$,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}. 
Under our scaling, $N$ and $\theta=\theta(N)$ will tend to infinity in such a 
way that $\alpha:=\lim_{N\to\infty}\theta(N)/N$ exists. From the 
expression~(\ref{eqn:prelimit_martingale_variation}) it is easy to guess that 
whether the limit points will be deterministic or stochastic processes 
is determined by whether $\alpha$ is zero or nonzero.
%c.f.~Section~\ref{sec:heuristics}.

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    We define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although sometimes we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align}
	    \nonumber
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    \label{drift in adjoint}
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        %\text{TODO} .
    \end{align}
\end{definition}

@@

We assume that 
\begin{enumerate}
\item The function $F(x,m)$ satisfies 
\begin{enumerate}
\item $F(x,m)$ is locally Lipschitz in $m$;
\item $F(x,m)$ is uniformly bounded above (but not necessarily below);
\item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
\end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
derivatives in both arguments;
\item $\gamma(x,m)$ is uniformly bounded above;
\item Either
\begin{enumerate}
\item
\label{control through r} 
$|\partial_{x_i}r(x,\smooth{r}\eta(x)|$ and 
$|\partial_{x_ix_j}r(x,\smooth{r}\eta(x)|$
are uniformly bounded for $x\in\IR^d$, $\eta\in\measures$;
\item 
\label{control through gamma}
or, $m^2\gamma(x,m)$ is uniformly bounded and there exists $C<\infty$
such that for $\theta$ sufficiently large, and all $x\in\IR^d$, $\eta\in\measures$,
\[
\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)q_\theta(x,dy)\leq C\smooth{\gamma}\eta(x),
\]
and
\[\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)^2q_\theta(x,dy)
\leq C(\smooth{\gamma}\eta(x))^2.
\]
\end{enumerate} 
\end{enumerate}

@@

The purpose of the conditions that we have placed on the reproduction parameters is to 
ensure that the net per capita reproduction rate is order $1/\theta$. As remarked above, 
because of the non-local reproduction mechanism, it no longer suffices to assume that 
$r(x,\smooth{r}\eta(x))\gamma(x, \smooth{\gamma}\eta(x))-\mu_\theta(x)$ is of order 
$1/\theta$. The conditions that we have imposed are to ensure that the difference 
between the establishment probability of a juvenile born of a parent at $x$ and dispersed
to a position $y$, determined by $q_{\theta}(x,dy)$, and its establishment probability had
it remained at $x$ are also order $1/\theta$.

If $r(x,m)$ is independent of $m$, then the conditions are easy to 
satisfy; they just require some regularity of $r$ as a function of $x$. 
Condition~\ref{control through r} is satisfied if 
$\|\nabla\rho_r\|\leq C\rho_r$ 
and $m\partial_mr(x,m)$, $m^2\partial_{mm}r(x,m)$ are bounded.
This will be satsified if $\rho_r$ decays 
no faster than
exponentially.
On the other hand, it might, for example, seem more natural to take $\rho_r$ to be a Gaussian 
density with parameter $\sigma_r$, say. Then, as we check in Lemma~\ref{old 6.1}, 
Condition~\ref{control through gamma} is satisfied if $\rho_\gamma$ is also
Gaussian with parameter $\sigma_\gamma$ and $\sigma_\gamma>\sigma_r$. For large enough
$\theta$, this condition guarantees that 
$\epsilon_r+1/\theta <\epsilon_\gamma$, so that the establishment probability of a juvenile
is controlled by individuals that are already `felt' by the fecundity kernel $\rho_\gamma$
at the location of their parent.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is tight 
in $\measures$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale,
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations \eqref{eqn:limiting_mgale_problem} and \eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover the superprocess with nonlinear death rates corresponding to logistic growth.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson Girsanov transform tells us that we have uniqueness in this special 
case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the spatial component of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that the martingale problem is well-posed in the case $\alpha>0$.

@@

\citet{KURTZ1999103} provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.

>>>>

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on space x level

@@

It is of interest to understand under what conditions we can replace the two-step limiting
process described above by one in which we simultaneously scale the 
kernels
and the other parameters in 
our population model to arrive at the PDE limit.
This is mathematically much more challenging, but we establish such 
one-step convergence in cases for which the limit is a classical 
reaction diffusion equation of the form~(\ref{semilinear PDE}) with 
$\DG=\Delta$,
and $\rho_F$ is a Gaussian density. We allow a wide class of reaction terms, $F$, so
that the Fisher-KPP equation emerges as a special case.

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity that have been laid down; from 
genetic data, one can try to infer the genealogical trees that relate 
individuals in a sample from the population, and these have been shaped 
by its history. It is therefore of interest
to establish information about the distribution of genealogical trees 
under our population model.
To achieve this we exploit a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a huge range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

We aim to capture universal behaviour by passing to a scaling limit. Specifically, 
we shall take the ``density'', $N$, to infinity,
and also scale time by a factor of $\theta = \theta(N)$,
in such a way that defining $\eta^N(t) = X(\theta t) / N$,
the process $\{\eta^N(t)\}_{t \ge 0}$
will converge to a suitable measure-valued process
as $N$ and $\theta$ tend to infinity,
with the nature of the limit depending on how they tend to infinity together.
Evidently, we shall also need to scale the dispersal kernel if we are to obtain a
nontrivial limit. We write $q_\theta(x,\cdot)$ for the density of the 
multivariate Gaussian obtained by
multiplying the mean and variance components corresponding to $q(x,\cdot)$ by $1/\theta$.

@@

\begin{remark}
Although this model allows fairly general birth and death mechanisms,
there are a number of limitations.
Perhaps most obviously, to simplify the notation
individuals give birth to only one offspring at a time,
although this restriction could be easily lifted
\citep[as in Section 3.4 of][]{etheridge/kurtz:2019}.
Furthermore, individuals do not move during their lifetime,
and the age of an individual does not affect its fecundity or death rate.
Finally, there is no notion of mating
(although limitations on reproduction due to availability of mates can be incorporated into the birth rate, $\gamma$),
so the lineages we follow will be uniparental.
For these reasons, the model is most obviously applicable to bacterial populations or selfing plants,
although we do not anticipate that incorporation of these complications
will change the general picture.
\end{remark}

@@

\begin{notation}
Expressions like $\gamma(x, \smooth{\gamma} \eta(x))$ will appear repeatedly in what follows.
To make formulae more readable, we overload notation to define
$$
    \gamma(x, \eta) := \gamma(x, \smooth{\gamma} \eta(x)) ,
$$
and similarly write $r(x, \eta)$ for $r(x, \smooth{r} \eta(x))$
and $\mu_\theta(x, \eta)$ for the expression of equation~\eqref{eqn:mu_defn}.
When convenient, we may also suppress the arguments completely,
writing simply $\gamma$, $r$, and $\mu_\theta$ for these quantities.
\end{notation}

@@

In summary, at each time $t$, $\eta^N_t$ is purely atomic, consisting of atoms of mass $1/N$.
At instantaneous rate $\theta \gamma(x, \eta^N_t) N \eta^N_t(dx)$ an offspring of mass $1/N$ is 
produced at location $x$; it disperses to 
a location $y$ offset from $x$
by an independent Gaussian random variable with mean $\meanq(x) / \theta$
and covariance matrix $\covq(x) / \theta$, and once there establishes instantaneously
with probability $r(y, \eta^N_t)$, or else dies. 
At instantaneous rate $\theta \mu_\theta(x, \eta^N_t) N \eta^N_t(dx)$ an individual at location 
$x$ dies.
Note that the process 
$\left(\eta^N_t\right)_{t\ge 0}$, which
records numbers and locations of adult individuals, is just a scaled spatial birth and 
death process. If, for example, we insist that $\gamma(x,m)$ is bounded, then 
existence (and in particular non-explosion) is guaranteed by comparison with
a pure birth process. We do not dwell on this, as we shall require more stringent conditions
if we are to pass to the limit as $\theta$ and $N$ tend to infinity.

It is convenient to characterise the process as a solution to a martingale problem.
We write $C_b^\infty(\IR^d)$ for the space of bounded smooth functions on $\IR^d$, and, where
convenient, we
write $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$.
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction}
	For each value of $N$ and $\theta$, and each purely atomic 
$\eta_0^N\in\measures$ with atoms of mass $1/N$,
$(\eta^N_t)_{t \geq 0}$ is the (scaled) empirical measure of a birth-death 
process with c\`adl\`ag paths in $\measures$ for which,
for all $f \in C^{\infty}_{b}(\IR^d)$,
writing $q_\theta(x, dy)$ for the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
%\\ & \qquad \qquad \qquad \qquad {}
         \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad \qquad \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale (with respect to the natural filtration), 
with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N(f)]_t =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \Big\langle \gamma(x, \eta^N_{s})
        \int f^2(z)r(z, \eta^N_{s}) q_\theta(x,dz) 
    , \eta^N_{s}(dx) \Big\rangle \\
& \qquad \qquad {}
    + \Big\langle \mu_\theta(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\Big\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}
The form of~(\ref{eqn:eta_martingale}) and~(\ref{eqn:prelimit_martingale_variation})
is explained in Section~\ref{sec:heuristics}.
Note that since (juvenile) individuals are produced at rate $N \gamma \eta$, but each
has mass $1/N$,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}. 
Under our scaling, $N$ and $\theta=\theta(N)$ will tend to infinity in such a 
way that $\alpha:=\lim_{N\to\infty}\theta(N)/N$ exists. From the 
expression~(\ref{eqn:prelimit_martingale_variation}) it is easy to guess that 
whether the limit points will be deterministic or stochastic processes 
is determined by whether $\alpha$ is zero or nonzero.
%c.f.~Section~\ref{sec:heuristics}.

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although sometimes we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align}
	    \nonumber
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    \label{drift in adjoint}
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        %\text{TODO} .
    \end{align}
\end{definition}

@@

We assume that 
\begin{enumerate}
\item The function $F(x,m)$ satisfies 
\begin{enumerate}
\item $F(x,m)$ is locally Lipschitz in $m$;
\item $F(x,m)$ is uniformly bounded above (but not necessarily below);
\item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
\end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
derivatives in both arguments;
\item $\gamma(x,m)$ is uniformly bounded;
\item Either
\begin{enumerate}
\item
\label{control through r} 
$|\partial_{x_i}r(x,\smooth{r}\eta(x)|$ and 
$|\partial_{x_ix_j}r(x,\smooth{r}\eta(x)|$
are uniformly bounded for $x\in\IR^d$, $\eta\in\measures$;
\item 
\label{control through gamma}
or, $m^2\gamma(x,m)$ is uniformly bounded and there exists $C<\infty$
such that for $\theta$ sufficiently large, and all $x\in\IR^d$, $\eta\in\measures$,
\[
\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)q_\theta(x,dy)\leq C\smooth{\gamma}\eta(x),
\]
and
\[\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)^2q_\theta(x,dy)
\leq C(\smooth{\gamma}\eta(x))^2.
\]
\end{enumerate} 
\end{enumerate}

@@

The purpose of the conditions that we have placed on the reproduction parameters is to 
ensure that the net per capita reproduction rate (before time scaling)
is order $1/\theta$. As remarked above, 
because of the non-local reproduction mechanism, it no longer suffices to assume that 
$r(x,\smooth{r}\eta(x))\gamma(x, \smooth{\gamma}\eta(x))-\mu_\theta(x)$ is of order 
$1/\theta$. The conditions that we have imposed are to ensure that the difference 
between the establishment probability of a juvenile born of a parent at $x$ and dispersed
to a position $y$, determined by $q_{\theta}(x,dy)$, and its establishment probability had
it remained at $x$ are also order $1/\theta$.

If $r(x,m)$ is independent of $m$, then the conditions are easy to 
satisfy; they just require some regularity of $r$ as a function of $x$. 
\comment{Not sure next sentence strictly true - probably assuming a bit
more - certainly some regularity.}
Condition~\ref{control through r} is satisfied if 
$\|\nabla\rho_r\|\leq C\rho_r$ 
and $m\partial_mr(x,m)$, $m^2\partial_{mm}r(x,m)$ are bounded.
This will be satsified if $\rho_r$ decays 
no faster than
exponentially.
On the other hand, it might, for example, seem more natural to take $\rho_r$ to be a Gaussian 
density with parameter $\sigma_r$, say. Then, 
as we check in Lemma~\ref{old 6.1}, \comment{this is referring to the
old Lemma 6.1, now to go in Appendix,}
Condition~\ref{control through gamma} is satisfied if $\rho_\gamma$ is also
Gaussian with parameter $\sigma_\gamma$ and $\sigma_\gamma>\sigma_r$. For large enough
$\theta$, this condition guarantees that 
$\sigma_r+1/\theta <\sigma_\gamma$, so that the establishment probability of a juvenile
is controlled by individuals that are already `felt' by the fecundity kernel $\rho_\gamma$
at the location of their parent.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
purely atomic measures, with atoms of mass $1/N$, which is
tight 
in $\measures$ and such that there exists $\lambda>0$ for 
which $\langle e^{\lambda\|x\|}, \eta_0^N(dx)\rangle<C<\infty$
with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover the superprocess with nonlinear death rates corresponding to logistic growth.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson Girsanov transform tells us that we have uniqueness in this special 
case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the 
dispersal mechanism of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that the martingale problem is well-posed in the case $\alpha>0$.

@@

Specialising the results of 
\citet{KURTZ1999103} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
